The human in the loop: volunteer computing as a socio-technical system

### 2016-02-29 Alife16/paper.tex
explaining a joke makes it less funny, but...
First part of experiments
@JJ please check some inconsistencies

In this case the classical Trap function \citep{Ackley1987} has been
used. JavaScript is a functional language and declaring a different 
function and handing at the creation of the algorithm object, called
{\tt Classic}, is the only change needed to work with a different
problem. The next Section will describe experiments performed to
establish baseline performance and gather initial performance
results. In the experiments performed last year we used the 40-trap
problem, while the current experiments changed to the more difficult 50-trap.

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rrl}
\hline
Experiment & \#Runs & Different IPs & Traps \\
\hline
April 4th 4/4 & 57 & 191 & 40 \\
April 24th 4/24 &  231 & 559& 40  \\
July 31th 7/31 & 97 & 179 & 40 \\
\hline
February, cache=128 & 61 & 75 & 50  \\
February, cache=64 & 61 & 220 & 50  \\
February, cache=32 & 39 & 86 & 50  \\
% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
\hline
\end{tabular}
\end{center}
\end{table}
%
Initial experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to 0. Experiments were
announced through a posts in Twitter and, the latest case, Telegram, and
results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of April and then by the end of February. All
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 40-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible.

@@

That is why we used a more difficult problem in the second batch of
experiments, which is shown in bottom row of Figure
\ref{fig:duration}. The pattern is remarkably similar, showing a
positive correlation between the time for solving the problem and the
number of computers, at least for cache sizes 128 and 64. However, it
is interesting to observe that, for cache=32, the time needed to find
the solution decreases from one to approximately 4-5 nodes, to then
increase for a higher number of participating computers, distinguished
by IP. The green dot at the bottom is probably and outlier that we
will try to explain later on. This leads us to conclude that a better
amount of computers might contribute to speed up the solution, if the
time the experiment ideally takes is sufficient, that is, of the order of a
minute, and enough volunteers concur simultaneously. This is also
observed, not so clearly, in the case of cache=64, with an interval of
around 10 IPs obtaining less time than experiments with less or more
IPs, and of the same order, between 10 and 100 seconds. If we look at
the graph that shows the number of IPs or volunteers per minute for
this experiment, shown in Figure \ref{fig:otisdriftwood}, we see that
there are peaks of more than 25 volunteers, and a period of several
hours with a minimum of 6 computers and peaks of more than 10. The
long period after midnight where there is a single volunteer left
masks the success achieved during this set of experiments, from which
we draw two lessons: first, you need a social network influencer to
announce your experiments and second, no matter what, do not do any
experiment after midnight.

>>>>

JavaScript is a functional language, so in order to work with a
problem, a fitness function must be supplied at the creation of the algorithm object, 
called {\tt Classic}. In this case the classical Trap 
function \citep{Ackley1987} has been used. Depending on the
problem additional handing functions and configurations 
can be supplied.

The next Section will describe experiments performed to
establish a baseline performance and gather initial performance
results. In the experiments performed last year we used the 40-trap
problem, while the current experiments changed to the more difficult 50-trap
in order to compare the performance with a more computationally intensive problem.

@@

\begin{table}[htb]
\caption{Experiment table, with summary of results. \label{tab:runs}}
\begin{center}
\begin{tabular}{l|rrl}
\hline
Experiment & \#Runs & Different IPs & Traps \\
\hline
April 4th 4/4 & 57 & 191 & 40 \\
April 24th 4/24 &  231 & 559& 40  \\
July 31th 7/31 & 97 & 179 & 40 \\
\hline
February, cache=128 & 61 & 75 & 50  \\
February, cache=64 & 61 & 220 & 50  \\
February, cache=32 & 39 & 86 & 50  \\
% grep start ~/Code/splash-volunteer/data/2016/nodio-2016-2-28-cache=32.log
\hline
\end{tabular}
\end{center}
\end{table}
%
Initial experiments were set up using the OpenShift
PaaS, which provides a free tier, making the whole experiment cost
equal to \$0.00. Experiments were
announced through a series of posts in Twitter and, in the latest case, Telegram, and
results were published here \citep{DBLP:conf/gecco/GuervosG15}. For the
purpose of this paper, we repeated the announcement several times
through the month of February. All  %% April --> January ??
in all, we have the set of runs with the characteristics shown in
Table \ref{tab:runs}. In general, every experiment took several
days. No particular care was taken about the time of the announcement
or the particular wording. Every {\em experiment} consisted in running
until the solution of the 50-trap problem was found. When the correct
solution was sent to the server, the counter was updated and the pool
of solutions reset to the void set. There was no special intention to wait
until all clients had finished, thus it might happen that, in fact,
the islands running in the browser {\em spill} from one experiment to
the next. However, previous experiments have proven that the influence
of these islands in the next experiment is indeed negligible.

@@

That is why we used a more difficult problem in the second batch of
experiments, which is shown in bottom row of Figure
\ref{fig:duration}. The pattern is remarkably similar, showing a
positive correlation between the time for solving the problem and the
number of computers, at least for cache sizes 128 and 64. However, it
is interesting to observe that, for cache=32, the time needed to find
the solution decreases from one to approximately 4-5 nodes, to then
increase for a higher number of participating computers, distinguished
by IP. The green dot at the bottom is probably and outlier that we
will try to explain later on. This leads us to conclude that a better
amount of computers might contribute to speed up the solution, if the
time the experiment ideally takes is sufficient, that is, of the order of a
minute, and enough volunteers concur simultaneously. This is also
observed, not so clearly, in the case of cache=64, with an interval of
around 10 IPs obtaining less time than experiments with less or more
IPs, and of the same order, between 10 and 100 seconds. If we look at
the graph that shows the number of IPs or volunteers per minute for
this experiment, shown in Figure \ref{fig:otisdriftwood}, we see that
there are peaks of more than 25 volunteers, and a period of several
hours with a minimum of 6 computers and peaks of more than 10. The
long period after midnight where there is a single volunteer left
masks the success achieved during this set of experiments, from which
we draw two lessons: first, you need a social network influencer to
announce your experiments and second, no matter what, do not do any
experiment after midnight. This statement, which might seem tongue in
cheek, in fact it is a conclusion drawn from the experimental data and
to what extent the social network is an essential part of the
description and performance of the NodIO volunteer computing system.
